Network Configuration Synthesis with Abstract Topologies \vspace{1cm}

### 2017-04-15 paper/abstraction/paper.tex
a bunch of grammar and spelling fixes

Unfortunately, templates use the same low-level constructs as ordinary router configurations (\EG,
adding or removing tags from particular announcements).  In both templatized
and non-templatized configurations, ensuring that a collection of low-level, local configurations
achieves some important global, network-wide policy, such as a guarantee of connectivity in the face of 
failures, is extremely challenging.
%
%Worse, operators can never be sure that their templates correctly implement the policy. %
%Formally verifying them is difficult; today's hand-generated templates have no particular semantics---which 
%policies they implement and for which topologies---other than those of the configurations obtained from 
%instantiating them.
%
In order to validate  their templates, operators will typically first instantiate a template with appropriate
concrete parameters
and then test it under various scenarios.  In general, like in any complex software system, such testing is inherently incomplete.
%Thus, operators validate templates by testing some concrete instantiations using
%simulation or verification tools~\cite{feamster+:rcc, batfish, era, arc, bagpipe}.

Moreover, even if network operators were to instantiate their templates using the initial network 
topology and verify key properties using tools such as Bagpipe~\cite{bagpipe}, the guarantees
would not hold as the network topology evolves. Evolution of the topology is a frequent event for large networks, as  devices and links are taken offline for maintenance and  added to expand
capacity.  
Templates that
work for the current topology may or may not work for future
topologies. Ensuing problems may cause operators to make non-uniform changes to particular routers' configurations,
which defeats the purpose of a template system.  An even worse situation is when
operators have to update many devices to evolve their network.  
Such network-wide configuration changes entail a great deal of risk and can be highly
disruptive to live traffic.
%Ideally, network engineers would like some guarantees about the future, namely that 
%certain types of changes to the network topology may be (1) made \emph{locally}, and do not require
%updates of routers far from the source of the change, and (2) do not violate key properties, such
%as fault tolerance.
%Updating the templates, if found to not work for a future topology,
%may cause trickle-down, network-wide configuration changes that are highly
%disruptive in production networks with live traffic.
%Building sophisticated tools for validating templates is difficult

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{narain+:configassure,propane,synet-synthesis}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, current synthesis systems operate over concrete topologies.  
Even if two devices play the same role, operators cannot specify policy in terms of
this role; and even if specifications for the two devices are similar,
there is no guarantee that the systems will generate (syntactically) similar
configurations.  Perhaps most importantly, if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of just a handful of
role configurations. Current synthesis systems are also brittle in the face of network
evolution. Any change in network topology requires re-execution of the engine,
from scratch, on the new topology, and the result may be a completely different
set of configurations.  No operator is
able to shut down a large, production network, upgrade policy on all devices and 
then restart their network.

@@

When announcing destinations to certain neighbors, a BGP router may perform {\em route aggregation}, \IE, announce a single prefix that covers multiple prefixes (\EG, announce 10.1.0.0/16 instead of 10.1.[1--255].0/24). The covering prefix is announced as long as the router has a valid route for {\em any} of the covered prefixes. Aggregation helps reduce the memory consumption of routers, but it can lead to traffic black holes~\cite{route-aggregation}. A router that announces 10.1.0.0/16, because it has a route to 10.1.1.0/24, may also get traffic for 10.1.2.0/24 to which it has no route.

@@

For example, in Figure~\ref{fig:example3}, the first constraint ($e1 = AG$) 
%and
states that, within any pod P, the number of outgoing edges from a node in the 
$TG$ role (\IE, $e1$) to a node in the $AG$ role equals the number of nodes in the 
$AG$ role. Similarly, the constraint ($e2 = TG$) states that the number of 
outgoing edges from a node in the $AG$ role (\IE, $e2$) to a node in the $TG$ role 
equals the number of nodes in the $TG$ role. Together these constraints capture 
the fact that, within any pod, the global aggregators and ToRs are in 
a full mesh. Furthermore, the constraints $AG = AL$ and $AG \leq S$ ensure 
that, within pods P and Q, the $AG$ and $AL$ roles \dpw{This originally said $AG$ and $TG$ but I changed it.} have the same number of 
routers, which is less than or equal to the number of routers in the spine 
role $S$.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Similar constraints appear for the local aggregator role. The constraint $2 \leq S \leq 4$ makes explicit the possibility for growth, for example, by growing the network from Figure~\ref{fig:example} to Figure~\ref{fig:example2}. In general, we need not bound the number of spine routers to admit more concrete topologies, potentially at the expense of analysis precision. We also include the constraints $(S \mod AG) = 0$, and $(S \mod AL) = 0$~simply to show that constraints do not have to be in the form of inequalities. Operators can use logical formulas from any theory supported by modern SMT solvers.

@@

Next we can capture constraint (4) that traffic for local prefixes must stay within in the data center:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The \CD{Local} policy adds the $\CD{\Always(\In)}$ constraint for each local prefix described by the template variable $\CD{\$LP}$. This constraint ensures that traffic follows a path that matches an internal role (\IE, in the data center) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each level in the data center.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the data center through a peer.

@@

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is particularly useful for data center topologies where routers in one tier of the data center often have a particular striping pattern with another tier.

@@

The algorithm was able to infer that there is at least 1 disjoint path to any spine router, and at least 2 disjoint paths to any TL router. In this case, the analysis is precise. There exists a concrete network, namely the data center from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint.
%The bound on the fault tolerance between ToR and spine routers can also be used to determine that an aggregation-induced black hole may arise at the data center boundary after a single failure in some concrete instantiation.

@@

Template generation uses community tags to record the state of the automata from the Product Graph in every BGP announcement and updates these tags whenever messages are passed between routers. This mechanism ensures that BGP only considers paths that are allowed by the \sysname policy (\IE, paths in the PG). Routers will allow advertisments that correspond to an edge in the PG and will block any advertisments that do not.

@@

If $m \leq n$, then advertisements received for node $m$ can safely be preferred over those received for node $n$ after accounting for the network-wide impact of the choice. For example in Figure~\ref{fig:example-compilation}, S1 can receive advertisements in two different contexts $(S1,1,0)$ and $(S1,0,1)$ and must choose between messages received in these different contexts. Advertisements are preferred in state $(S1,1,0)$ because they result in a better path for S1 ($(S1,0,1) \geq_{rank} (S1,1,0)$) and downstream routers such as A1 will also obtain paths no worse than if S1 had choosen $(S1,1,0)$. The policy is guaranteed to be safe from failures because if a link fails in the topology, then $m \leq n$ will still hold since any transition that becomes unusable for $m$ also becomes unusable for $n$. That is $m \leq n$ before the failure implies $m \leq n$ after the failure.

@@

Figure~\ref{fig:bgp-configs} (left) shows part of the generated \mbgp configuration for spine routers for both the concrete and abstract policies. For brevity, we using the symbol ($*$) to denote the set of all neighbors and omit tags when irrelevant.
%
For prefix $\True$, the spine routers will match advertisements from peer N1 and N2. The match for N1 is preferred since it has a higher BGP local-preference attribute (110). If an advertisement from N1 is choosen, the spine attaches the community tag $(1,0)$ before sending the route to all its peers ($*$). If an advertisement is only available from the backup N2, then it attaches the tag $(0,1)$ instead.
%
The template configuration matches any global prefix {\small \$GP} from any internal peer and re-advertises the route to all of its peers. For any local prefix it will allow an advertisement from any internal peer, and re-advertise the route to only other internal peers. The concrete configurations for S1 and S2 obtained from compilation for the concerete PG from Figure~\ref{fig:example-compilation} have a similar structure for each local and global prefix where local routes are reflected downward, while global routes are advertised to all peers.

@@

Although the disjoint path analysis takes place over the PG, each application of the inference rules from Figure~\ref{fig:inference-rules} depends on the topology locations, but not the automata states, and can be reused accross multiple PG nodes with the same topology location. Therefore, we lazily apply the rules and cache the satisfiability and minimization calls to $\nu$Z after their first use. Furthermore, the cached results are shared across different prefixes, each of which may have a unique PG representation.

@@

The backbone policy classifies neighbors into several categories based on commercial relationship~\cite{gao-rexford} and prefers paths through them in order. Similar to the datacenter, it blocks private destinations from neighbors, drops transit traffic between certain neighbors but not others, and aggregates internal prefixes at the border of the network.
%
We scale the backbone network from 10 to 240 routers. To abstract it, we split it into two parts: border routers that connect to external neighbors and an internal core. We use one abstract node for the border routers and one for the network core with mincut annotations both within the core and between the core and border roles. For neighbors, there is one abstract role per commercial category.

>>>>

Unfortunately, templates use the same low-level constructs as ordinary router configurations (\EG,
adding or removing tags from announcements).  In both templatized
and non-templatized configurations, ensuring that a collection of low-level, local configurations
achieves some important network-wide policy, such as a guarantee of connectivity in the face of 
failures, is extremely challenging.
%
%Worse, operators can never be sure that their templates correctly implement the policy. %
%Formally verifying them is difficult; today's hand-generated templates have no particular semantics---which 
%policies they implement and for which topologies---other than those of the configurations obtained from 
%instantiating them.
%
To validate their templates, operators will typically first instantiate a template with appropriate
concrete parameters
and then test it under various scenarios.  In general, like in any complex software system, such testing is inherently incomplete.
%Thus, operators validate templates by testing some concrete instantiations using
%simulation or verification tools~\cite{feamster+:rcc, batfish, era, arc, bagpipe}.

Moreover, even if network operators were to instantiate their templates using the initial network 
topology and verify key properties using tools such as Bagpipe~\cite{bagpipe}, the guarantees
would not hold as the network topology evolves. Evolution of the topology is a frequent event for large networks, as  devices and links are taken offline for maintenance and  added to expand
capacity.  
Templates that
work for the current topology may or may not work for future
topologies. Ensuing problems may cause operators to make non-uniform changes to routers' configurations,
which defeats the purpose of a template system.  An even worse situation is when
operators must update many devices to evolve their network.  
Such network-wide configuration changes entail a great deal of risk and can be highly
disruptive to live traffic.
%Ideally, network engineers would like some guarantees about the future, namely that 
%certain types of changes to the network topology may be (1) made \emph{locally}, and do not require
%updates of routers far from the source of the change, and (2) do not violate key properties, such
%as fault tolerance.
%Updating the templates, if found to not work for a future topology,
%may cause trickle-down, network-wide configuration changes that are highly
%disruptive in production networks with live traffic.
%Building sophisticated tools for validating templates is difficult

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{narain+:configassure,propane,synet-synthesis}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, current synthesis systems operate over concrete topologies.  
Even if two devices play the same role, operators cannot specify policy in terms of
this role; and even if specifications for the two devices are similar,
there is no guarantee that the systems will generate (syntactically) similar
configurations.  Perhaps most importantly, if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of just a handful of
role configurations. Current synthesis systems are also brittle in the face of network
evolution. Any change in network topology requires re-execution of the engine,
from scratch, on the new topology, and the result may be a completely different
set of configurations.  No operator can shut down a large, production network, upgrade policy on all devices and 
then restart their network.

@@

When announcing destinations to certain neighbors, a BGP router may perform {\em route aggregation}, \IE, announce a single prefix that covers multiple prefixes (\EG, announce 10.1.0.0/16 instead of 10.1.[1--255].0/24). The covering prefix is announced if the router has a valid route for {\em any} of the covered prefixes. Aggregation helps reduce the memory consumption of routers, but it can lead to traffic black holes~\cite{route-aggregation}. A router that announces 10.1.0.0/16, because it has a route to 10.1.1.0/24, may also get traffic for 10.1.2.0/24 to which it has no route.

@@

For example, in Figure~\ref{fig:example3}, the first constraint ($e1 = AG$) 
%and
states that, within any pod P, the number of outgoing edges from a node in the 
$TG$ role (\IE, $e1$) to a node in the $AG$ role equals the number of nodes in the 
$AG$ role. Similarly, the constraint ($e2 = TG$) states that the number of 
outgoing edges from a node in the $AG$ role (\IE, $e2$) to a node in the $TG$ role 
equals the number of nodes in the $TG$ role. Together these constraints capture 
the fact that, within any pod, the global aggregators and ToRs are in 
a full mesh. Furthermore, the constraints $AG = AL$ and $AG \leq S$ ensure 
that, within pods P and Q, the $AG$ and $AL$ roles have the same number of 
routers, which is less than or equal to the number of routers in the spine 
role $S$.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Similar constraints appear for the local aggregator role. The constraint $2 \leq S \leq 4$ makes explicit the possibility for growth, for example, by growing the network from Figure~\ref{fig:example} to Figure~\ref{fig:example2}. In general, we need not bound the number of spine routers to admit more concrete topologies, potentially at the expense of analysis precision. We also include the constraints $(S \mod AG) = 0$, and $(S \mod AL) = 0$~simply to show that constraints do not have to be in the form of inequalities. Operators can use logical formulas from any theory supported by modern SMT solvers.

@@

Next, we can capture constraint (4) that traffic for local prefixes must stay within in the data center:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The \CD{Local} policy adds the $\CD{\Always(\In)}$ constraint for each local prefix described by the template variable $\CD{\$LP}$. This constraint ensures that traffic follows a path that matches an internal role (\IE, in the data center) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each level in the data center.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the data center through a peer.

@@

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is particularly useful for data center topologies where routers in one tier of the data center often have a uniform striping pattern with another tier.

@@

The algorithm could infer that there is at least 1 disjoint path to any spine router, and at least 2 disjoint paths to any TL router. In this case, the analysis is precise. There exists a concrete network, namely the data center from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint.
%The bound on the fault tolerance between ToR and spine routers can also be used to determine that an aggregation-induced black hole may arise at the data center boundary after a single failure in some concrete instantiation.

@@

Template generation uses community tags to record the state of the automata from the Product Graph in every BGP announcement and updates these tags whenever messages are passed between routers. This mechanism ensures that BGP only considers paths that are allowed by the \sysname policy (\IE, paths in the PG). Routers will allow advertisements that correspond to an edge in the PG and will block any advertisements that do not.

@@

If $m \leq n$, then advertisements received for node $m$ can safely be preferred over those received for node $n$ after accounting for the network-wide impact of the choice. For example, in Figure~\ref{fig:example-compilation}, S1 can receive advertisements in two different contexts $(S1,1,0)$ and $(S1,0,1)$ and must choose between messages received in these different contexts. Advertisements are preferred in state $(S1,1,0)$ because they result in a better path for S1 ($(S1,0,1) \geq_{rank} (S1,1,0)$) and downstream routers such as A1 will also obtain paths no worse than if S1 had chosen $(S1,1,0)$. The policy is guaranteed to be safe from failures because if a link fails in the topology, then $m \leq n$ will still hold since any transition that becomes unusable for $m$ also becomes unusable for $n$. That is $m \leq n$ before the failure implies $m \leq n$ after the failure.

@@

Figure~\ref{fig:bgp-configs} (left) shows part of the generated \mbgp configuration for spine routers for both the concrete and abstract policies. For brevity, we using the symbol ($*$) to denote the set of all neighbors and omit tags when irrelevant.
%
For prefix $\True$, the spine routers will match advertisements from peer N1 and N2. The match for N1 is preferred since it has a higher BGP local-preference attribute (110). If an advertisement from N1 is chosen, the spine attaches the community tag $(1,0)$ before sending the route to all its peers ($*$). If an advertisement is only available from the backup N2, then it attaches the tag $(0,1)$ instead.
%
The template configuration matches any global prefix {\small \$GP} from any internal peer and re-advertises the route to all its peers. For any local prefix, it will allow an advertisement from any internal peer, and re-advertise the route to only other internal peers. The concrete configurations for S1 and S2 obtained from compilation for the concrete PG from Figure~\ref{fig:example-compilation} have a similar structure for each local and global prefix where local routes are reflected downward, while global routes are advertised to all peers.

@@

Although the disjoint path analysis takes place over the PG, each application of the inference rules from Figure~\ref{fig:inference-rules} depends on the topology locations, but not the automata states, and can be reused across multiple PG nodes with the same topology location. Therefore, we lazily apply the rules and cache the satisfiability and minimization calls to $\nu$Z after their first use. Furthermore, the cached results are shared across different prefixes, each of which may have a unique PG representation.

@@

The backbone policy classifies neighbors into several categories based on commercial relationship~\cite{gao-rexford} and prefers paths through them in order. Like the data center, it blocks private destinations from neighbors, drops transit traffic between certain neighbors but not others, and aggregates internal prefixes at the border of the network.
%
We scale the backbone network from 10 to 240 routers. To abstract it, we split it into two parts: border routers that connect to external neighbors and an internal core. We use one abstract node for the border routers and one for the network core with mincut annotations both within the core and between the core and border roles. For neighbors, there is one abstract role per commercial category.
